---
title: "CaseStudy1"
author: "Jason Cessna"
date: "3/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##INTRODUCTION

This is a statistical analysis conducted on the Gross Domestic Product of countries in 2012 in direct relation to Education Data we have received which denotes the income level of countries. We will combine these data together to compare the relationship between the ranking of countries in terms of their GDP (gross domestic product) and their income grouping levels. To perform this analysis we will use R. The data files we will be using for analysis can be found here:
Gross Domestic Product of 2012:
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv
Eudcational Data:
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv

We will start our analysis by installing the proper packages into R (ggplot2, dplyr, and doBy). We will need these for our statistical analysis and visualization of our data.

##INSTALL PACKAGES
(the installation is currently commented out of this markdown as the mirrors have not been provided by the developer. find the mirrors for your country to download the packages here: https://cran.r-project.org/mirrors.html)
```{r}
##source('Analysis/installPackages.R', echo=TRUE)
```


```
##install.packages('ggplot2',repos = "http://cran.us.r-project.org")
##install.packages('doBy',repos = "http://cran.us.r-project.org")
##install.packages('dplyr',repos = "http://cran.us.r-project.org")
```

Then we download the data

##DOWNLOAD DATA
```{r visible=TRUE}
source('Analysis/downloadData.R', echo=TRUE)
```

Now we need to clean our data to make sure there is no unncessary blank spaces that would interfere with our analysis. We are also preparing to merge the two data files on their short code (or country code) so we have all relevant information for each country provided in both documents.

##CLEAN DATA

```{r}
source('Analysis/cleanData.R', echo=TRUE)
```
Now that our data is clean, we need to merge the two data frames together. In order to do so, we must make sure that the numerical data from our gross domestic product data frame is converted properly so the statistical analysis can be executed properly after the merge.

##MERGE DATA

```{r}
source('Analysis/mergeData.R', echo=TRUE)
```
Now that our data is merged, we can perform statistical analysis on it. The questions we are looking to answer are as follows:

## QUESTION 1
Merge the data based on the country shortcode. How many of the IDs match?
Since we already merged the data based on the country shortcode, we are just looking to discover how many matches we had in total (ie: how many countries have valid data for our analysis?)
```{r}
source('Analysis/question1.R', echo=TRUE)
```
Out of the 190 country codes in our Gross Domestic Product data and the 234 country codes in the Educational Data, we are able to match and merge on 189 total country codes.

## QUESTION 2
Sort the data frame in ascending order by GDP (so United States is last). What is the 13th
country in the resulting data frame?
```{r}
source('Analysis/question2.R', echo=TRUE)
```
The country with the 13th-lowest generated domestic product is St. Kitts and Nevis

## QUESTION 3
What are the average GDP rankings for the "High income: OECD" and "High income:
nonOECD" groups?
```{r}
source('Analysis/question3.R', echo=TRUE)
```
The average (or mean) of the GDP of countries in the High Income: OECD income group is $32.97 (in millions USD), and the mean of the GDP of countries in the High Income: nonOECD income group is $91.91 (in millions USD).

## QUESTION 4
Show the distribution of GDP value for all the countries and color plots by income group. Use
ggplot2 to create your plot.
```{r}
source('Analysis/question4.R', echo=TRUE)
```

## QUESTION 5
Provide summary statistics of GDP by income groups.
```{r}
source('Analysis/question5.R', echo=TRUE)
```
Below we have the summary statistics for each income group based on their GDP:
```
IncomeGroup             Min.    1st Qu.   Median    Mean      3rd Qu.    Max
High income: nonOECD    2584    12840     28370     104300    131200     711000 
High income: OECD       13580   211100    486500    1484000   1480000    16240000 
Low income              596     3814      7843      14410     17200      116400 
Lower middle income     40      2549      24270     256700    81450      8227000 
Upper middle income     228     9613      42940     231800    205800     2253000 
```
Below we have some additonal statistical data from each income group based on its GDP:
```
INCOME GROUP             MEAN      standard deviation(GDP)    sum of GDP     count of countries 
High income: nonOECD     104349.83        165334.45           2400046        23
High income: OECD        1483917.13       3070463.52          44517514       30
Low income               14410.78         20473.09            533199         37
Lower middle income      256663.48        1139619.92          13859828       54
Upper middle income      231847.84        476872.04           10433153       45
```
## QUESTION 6
Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group.
How many countries are Lower middle income but among the 38 nations with highest
GDP?
```{r}
source('Analysis/question6.R', echo=TRUE)
```
There are 16 countries in both the "lower middle income"" income group and the first quantile (top 38 countries) for Gross Domestic Product produced in 2012. This begs the question: how are these countries utilyzing their internal resources in terms of the betterment of their economy? What sorts of resources are produced to make up their GDP? What is the pay which the residents within that country receive while working to produce these resources (if any)? Our data alone cannot answer these questions, but we do have grounds for further investigation based on our results.

The table can be seen below:
```
Quantile | High income: nonOECD | High income: OECD | Low income |Lower middle income | Upper middle income
1                    2                    0                11               16                  9
2                    4                    1                16                9                  8
3                    8                    1                 9               11                  9
4                    5                   10                 1               14                  8
5                    4                   18                 0                4                 11
```

##CONCLUSION

The analysis we have conducted on these data has showed us a qualitative comparison between countries based on their income groups, rankings, and gross domestic product. We did not find that there is a constant correlation between a high GDP and high income within countries and raises more questions that we can perform additional analysis on if we can obtain more data (ie: products produced, services sold, mean wages for citizens in those jobs, and mean populations for each country) like the correlation between population's wages of a certain industry and the overall profits which the country receives on them.

We also see (in our plot) that only 3 countries have a GDP over 5 trillion in 2012, one of which had an income that places that country in the lower-middle income group. Less than 25% of countries had a GDP of above 1 trillion USD in 2012, and their incomes vary tremendously as well.

With both of these observations, we can conclude that there is not a direct correlation between the gross domestic product of a country and their overall income.


